\ProvidesFile{template.tex}
    [2013/04/16 v3.2013^^J%
     Template file for NDdiss2e class by Sameer Vijay and updated by Megan Patnott^^J]
\documentclass[noinfo,final,sort&compress]{nddiss2e}
% One of the options draft, review, final must be chosen.
% One of the options textrefs or numrefs should be chosen
% to specify if you want numerical or ``author-date''
% style citations.
% Other available options are:
% 10pt/11pt/12pt (available with draft only)
% twoadvisors
% noinfo (should be used when you compile the final time
%         for formal submission)
% sort (sorts multiple citations in the order that they're
%       listed in the bibliography)
% compress (compresses numerical citations, e.g. [1,2,3]
%           becomes [1-3]; has no effect when used with
%           the textrefs option)
% sort&compress (sorts and compresses numerical citations;
%           is identical to sort when used with textrefs)

% \usepackage{amsmath}
% % \usepackage{amssymb}
% \usepackage{graphicx}
% \usepackage{multirow}
% \usepackage{dcolumn}
%\usepackage{bm} 
% \usepackage{amssymb}
%\usepackage{tablefootnote}
% \usepackage{times}
% \usepackage{mathptm}
%\usepackage{caption}
% \usepackage{float}
% \usepackage{geometry}
% \usepackage{xkeyval}
% \usepackage{tabularx}
% \usepackage{longtable}
% \usepackage{multicol}
% \usepackage{square}{natbib}
% \usepackage{achemso}
% \usepackage{cleveref}
%----my packs--
\usepackage{tikz}
\usetikzlibrary{shapes,arrows}
\usepackage{pdflscape}
\usepackage[version=3]{mhchem}
\usepackage{mathtools}
\usepackage{abraces}
\usepackage{threeparttable}
\usepackage{multirow}
\usepackage{dcolumn}
\usepackage[export]{adjustbox}
\usepackage{mathrsfs}
\usepackage{ragged2e}
\usepackage{csquotes}
%----my packs--

% \usepackage{tikz}
% \usetikzlibrary{shapes,arrows}
% \usepackage{pdflscape}
% \usepackage[version=3]{mhchem}
% \usepackage{mathtools}
% \usepackage{abraces}
% \usepackage{threeparttable}

%\setcitestyle{square}

%\newcommand*{\citen}[1]{%
 % \begingroup
  %  \romannumeral-`\x % remove space at the beginning of \setcitestyle
   % \setcitestyle{numbers}%
   % \cite{#1}%
  %\endgroup   
%}

\newcommand{\degree}{\ensuremath{^\circ}}

\begin{document}

\frontmatter % All the items before the first chapter go in ``frontmatter''

\title{DISCOVERING THE MOLECULAR ORIGINS OF SOLID-LIQUID FRICTION AT 
  ICE-I$_\mathrm{h}$/WATER INTERFACES}
\author{Patrick B. Louden}
\work{Dissertation}
\degaward{Doctor of Philosophy}
\advisor{J. Daniel Gezelter}
\department{Chemistry and Biochemstry}
\maketitle
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		FRONT STUFF
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%\copyrightholder{Patrick B. Louden} 
%\copyrightyear{2017}           
%\makecopyright

\begin{abstract}
  In this dissertation I present work on the modeling of hydrodynamic
  friction at the surface of ice. In the hydrodynamic regime, a slider
  traversing an ice surface is fully supported by a
  quasi-liquid layer (QLL), and the resulting friciton is comprised of
  three parts; solid-liquid friction between the ice crystal and the
  QLL, viscous shearing of the surface premelt, and the drag force due
  to capillary bridges forming between the QLL and the slider.

  Using non-equilibrium molecular dynamics simulations, the basal
  $\{0001\}$, prismatic $\{10\bar{1}0\}$, 14\degree~pyramidal
  $\{20\bar{2}1\}$, and secondary prismatic $\{11\bar{2}0\}$ facets of
  ice-I$_\mathrm{h}$ were drawn through liquid water with a momentum
  flux between the solid and liquid phases. The spatial transition
  between the ice and the liquid was quantified using structural and
  dynamic order parameters. Estimates of the interfacial width using
  structural measures gave widths of $\sim$ 6-10~\AA. Dynamic
  interfacial widths were quantified and predicted a slightly broader
  interface, $\sim$~10-15~\AA~ wide. In addition, spatial
  decomposition of molecular orientational correlation functions
  indicated that the short- and longer-time decay components behave
  differently closer to the interface. In all cases the interface was
  observed to be stable under the presence of a shear.

  Water sliding over the surface of ice is observed to be in the
  no-slip limit. A friction coefficient appropriate for negative slip
  boundary conditions is presented, and crystal facet dependent
  friction is observed. The computed friction of these interfaces is
  found to be invariant to the shear rate and direction of shear
  relative to the surface features. This trend is observed for two
  water potentials, and is shown to correlate strongly with the
  surface hydrogen-bond density. A momentum transmission model is
  proposed, which relates the observed friction with the density of
  solid to liquid hydrogen bonds, the shear viscosity of the liquid,
  and the width of the interface. Data collected from both models
  agrees well with the proposed model.

  A temperature dependence study of the surface premelt at the basal
  and prismatic facets of an ice-I$_\mathrm{h}$ crystal is
  presented. Spatially resolved diffusion constants are computed
  normal to the ice surface, and estimates of the shear viscosity for
  the QLLs are obtained. A facet depedence is observed in the shear
  viscosity which correlates strongly with the computed solid-liquid
  friction coefficients. Structural and dynamic properties of the QLLs
  are compared with simulations of supercooled liquid water. It is
  observed that the QLLs have characteristics distinctly different
  than the supercooled bulk liquid, attributed to the presence of the
  vapor interface.
\end{abstract}

% \renewcommand{\dedicationname}{NEW DEDICATION NAME}
% \begin{dedication}
% \end{dedication}

\tableofcontents
\listoffigures
\listoftables

\begin{preface}
\vfill
\begin{flushright}
\textit{``Some say the world will end in fire,} \\
\textit{Some say in ice.''}\\
--Robert Frost
\end{flushright}
\vfill
\end{preface}

\begin{acknowledge}
  This document is the culmination of the work and support of many
  wonderful educators, mentors, colleagues, and friends. I would like
  to especially thank my advisor, Professor J. Daniel Gezelter, for
  his countless hours of explanation, teaching, and
  encouragement. Malcolm Gladwell has estimated that it takes 10,000
  hours of practice and experience to gain expertise in any field. I
  have no doubt that I have asked Dan at least 10,000 questions in my
  time at Notre Dame, and with unwaivering patience he has answered
  every one. I am grateful to past and present group members,
  Dr. Shenyu Kuang, Dr. James Marr, Dr. Kelsey Stocker, Dr. Madan
  Lamichane, Suzie Neidhart, Hemanta Bhattarai, and Alex Mazanek. I am
  deeply grateful to Dr. Joseph Michalka for all the years of guidance
  and mentorship in programming and discussions of theoretical
  chemistry.

  I am thankful to the Corcelli and Maginn groups for their useful
  discussions and camaraderie throughout the tough times of
  research. Scientific discovery and understanding are not always
  straight paths, and I am grateful for the companions who made the
  journey with me. I also am thankful for the excellent courses taught by
  Professor Steven Corcelli, Professor Jindal Shah, Professor Bruce
  Bunker, Professor Kathie Newman, and Professor J. Daniel Gezelter.

  I am very thankful to my committee members, Professor J. Daniel
  Gezelter, Professor Steven Corcelli, and Professor Edward Maginn,
  for their guidance, wisdom, and helpful suggestions; all of which
  have helped make this document possible.

  I am also grateful to my undergraduate research adviser and close
  friend, Professor Christopher P. Lawrence. If I have achieved
  anything in my graduate research, it was only made possible by the
  foundation that he laid.

  I would be remiss if I did not acknowledge my close friends in
  chemistry, Dr. Robert Graff, Dr. Eric Hansen, and Tiffany Snow, as
  well as my extended family in physics, James Kelly, Anthony Ruth,
  Matthew Hall, Alison Hall, and Alison Deatsch. 

  I would like to thank my parents, for without whom I would not be
  where I am today. All of your years of sacrifice and encouragement
  have contributed to this document, and I am forever thankful for
  your love and support.

  Finally, I would like to acknowledge and thank my best friend and
  wife, Elizabeth Louden. Her continued support and understanding
  during my time at Notre Dame has meant the world to me, especially
  while writing this document. Her ability to make me smile continues
  to astound me.

  % Computational resources were provided by the University of Notre
  % Dame Center for Research Computing. Financial support for this
  % research was provided by the National Science Foundation under Grant
  % CHE-0848243 and the University of Notre Dame Department of Chemistry
  % and Biochemistry.
\end{acknowledge}

\mainmatter

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CHAPTER 1 -- INTRODUCTION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{Chapters/introduction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CHAPTER 2 -- METHODS
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{Chapters/iceWater}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CHAPTER 3 -- STRUCTURAL MEASURES OF INT. WIDTH
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{Chapters/structuralWidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CHAPTER 4 -- DYNAMIC MEASURES OF INT. WIDTH
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{Chapters/dynamicWidth}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CHAPTER 5 -- FRICTION AT ICE / WATER INTERFACES
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{Chapters/friction}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CHAPTER 6 -- QLL VISCOSITY
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{Chapters/qllViscosity}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		CHAPTER 7 -- CONCLUSION
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\include{Chapters/conclusion}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		UNNUMBERED CHAPTER
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% \unnumchapter{FEATURES OF FORMATTING IN THIS EXAMPLE FILE}
% The \unnumchapter command allows you to include an unnumbered chapter as part of
% the main text before Chapter 1. It will appear in your table of contents, and you
% should have at most one such chapter (although nothing in the class file will
% prevent you from creating more).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		APPENDIX
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
% \include{appendix}
%\include{Chapters/appendix1}
\include{Chapters/structuralAppendix}
\include{Chapters/dynamicAppendix}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%		BACK STUFF
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% % comment out the following three lines
% if using chapter-wise bibliography

 \backmatter
 
 % \bibliographystyle{nddiss2e}
 % \bibliographystyle{nddiss2enoarticletitles}
 \bibliographystyle{achemso}
 % \bibliographystyle{abbrvnat} 
  %\bibliography{Stocker_Thesis} 
 \bibliography{Louden_Thesis} 
 
 % The standard abbrvnat style should be acceptable. Also provided with both the advanced and standard distributions are nddiss2e and nddiss2enoarticletitles style options. If you prefer to manually enter your bibliography, that is fine. Comment out the previous two lines, and enter your bibliography as usual. Note that if you choose this route, formatting the bibliography is your responsibility. An example is below, including the optional arguments necessary for author-date style citations.
 
%	\begin{thebibliography}{9}
%		\bibitem[Galmira(1998)]{galmira98:_gnus_milit} G.\ Galmira. Gnus and the military -- a secret conspiracy? \emph{Growing Towards Gnu}, III(7):22--183, September 1998.
%		
%		\bibitem[Ganston and Greenfield(1998)]{gnus98:_gerry_ganst} G.\ Ganston and G.\ Greenfield. \emph{Gnus and You: The Art of Being New}. volume I. Grapping Books, NY, August, 1998.
%		
%		\bibitem[Gloonston(1998)]{gloonston98:_gnuly_discov_gnus} G.\ Gloonston. Newly discovered gnus: The LoG. \emph{Growing Towards Gnu}, II(12):23---57, March 1998.
%		
%		\bibitem[Greenfield(1996)]{greenfield96:_gettin_know_gnu} G.\ Greenfield. \emph{Getting to Know Gnu}. PhD thesis, Geoffrey Garfield School of Gnus, August 1996.
%		
%		\bibitem[van Gairley(2000)]{gairley2000} G.\ van Gairley. Gnu's review. Website, 2000. \url{http://www.gairley.gnu}.
%	\end{thebibliography}

\end{document}

\endinput